---
title: "Strings and Factors"
output: github_document
---
```{r}
library(tidyverse)
library(p8105.datasets)
library(rvest)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

## all plots i make will have the viridis color palette
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


## Strings and regex

Create strings as vectors 
detects a pattern in the string . We want to detect where "Jeff" is
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")

str_replace(string_vec, "jeff", "Jeff")


```

"^i think" detect strings that begin with "i think"
"i think$" detects strings that end with "i think"
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")

str_detect(string_vec, "i think$")
```


"[Bb]" gives both strings with B or b in the string
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )
str_detect(string_vec, "bush")
str_detect(string_vec,"[Bb]ush")
```


[0-9] anything between those 
[a-zA-Z] anything betwen a-z capital or lowercase
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")

```


"." matches anything in the string (dot is a special character)
to match a literal dot, you have tell it that it is not a special character ("\\.")
```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
str_detect(string_vec, "7\\.11")
 
```


detect things NOT AS special characters
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## Factors

Create a vector with a factor 
"Levels" 
By converting it to a numeric, we can see the integer baseline that underlies the factor
```{r}
factor_vec = factor(c("male", "male", "female", "female"))
factor_vec

as.numeric(factor_vec)
```

What happens if I relevel?
Noe male is listed as 1 instead of 2 
There are lots of fct functions
```{r}
factor_vec = fct_relevel(factor_vec, "male")

as.numeric(factor_vec)
```


## NSDUH -- strings


```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

tbl_marj = 
  read_html(url) %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>%
  slice(-1) %>% 
  as_tibble()
```

Want to clean up tibble
"(-contains)" removed everything that contains the following
Pivot everything that isn't in the State column
Easiest way to separate is to find the open parentheses (parentheses is a special character so we have to add "\\")
"str_replace" to get rid of ending ")" in year variable 
"!State" don't want state variable to be ....
```{r}
data_marj = 
  tbl_marj %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  ) %>% 
  filter((!State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


## NSDUH -- factors

can use fct_relevel in mutate statement to list out state you want to appear first in the graph
fct_reorder reorders the data based on the percent variable
```{r}
data_marj %>% 
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year))+
  geom_point() +
  theme (axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```













